---
title: "INA General Examples"
author: "Theodore Stronkowsky IV"
date: "2024-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Useful/Necessary Libraries


```{r}
library(ggplot2)
library(ggraph)
library(tidygraph)
library(tidyr)
library(tidyselect)
library(tidyverse)
library(igraph)
library(grid)
library(graphlayouts)
library(geodata)
library(INA)
```


## Cleaning Data and genlocs function

Generates longitudinal and latitudinal coordinates for each node.


```{r}

# Set seed for reproducibility
set.seed(42)

# Step 0: Data Cleaning (NEEDS to be done before using INA functions)
# This would include checking for NA values, duplicates, ensuring proper data types,
# and normalizing data if necessary.

# Step 1: Generate random geographic locations of nodes
genlocs_data <- genlocs(xrange4=c(0,50), yrange4=c(0,50), numnodes4=100, randgeo4=TRUE)
```

## initsvals function
Determine intial status of threat at node based on pest extent and managment preactice

```{r}
# Step 2: Initialize the network with random values
initvals_data <- initvals(genlocs_data, dist4='random', init.p4=0.1, norp4='prop', plotmp=TRUE)
```

## genmovnet function
Generates adjacency matrix for movement.

```{r}
# Step 3: Generate a random movement network based on the locations
# Needs work refining the network
genmovnet_data <- genmovnet(genlocs_data, amdist4='random', amrandp4=0.01, iplot=TRUE)
```

## setup2 function
Set up the initial conditions for INA scenario analysis and multisame2


```{r}
# Step 4: Setup starting conditions for INA scenario analysis
setup2_data <- setup2(
  maneffmean3s = 0.5,
  maneffsd3s = 0.5,
  maneffthresh3 = 0.5,
  sampeffort3 = 1,
  usethreshman3 = TRUE,
  readgeocoords3s = FALSE,
  geocoords3s = NULL, 
  xrange3 = c(0,50),
  yrange3 = c(0,50),
  numnodes3 = 100,
  randgeo3 = TRUE,
  readinitinfo3 = FALSE,
  initinfo3 = NULL, 
  initinfo.dist3 = 'random',
  initinfo.n3 = 5,
  initinfo.norp3 = 'num',
  initinfo.p3 = NA, 
  readinitbio3 = FALSE,
  initbio3 = NULL, 
  initbio.dist3 = 'random',
  initbio.n3 = 5,
  initbio.norp3 = 'num',
  initbio.p3 = NA, 
  plotmp = TRUE
)
```
## estab function
Estimates effects of managment practice on pest extent/establishment based on the initial status and locations


```{r}
# Step 5: Estimate the establishment probability at each node
estab_data <- estab(
  decvec = rep(1, nrow(genlocs_data)), 
  dispvec = rep(1, nrow(genlocs_data)), 
  geocoords4n = genlocs_data,  # Here we pass the matrix directly
  readprobestabvec4 = FALSE, 
  probestabmean4 = 0.5, 
  probestabsd4 = 0.1, 
  maneffdir4 = 'decrease_estab', 
  maneffmean4n = 0.1, 
  maneffsd4n = 0.1, 
  plotmp = TRUE
)
```
## onestart function
Evaluate individual nodes for invasion detection value

```{r}
# Step 6: Simulate an invasion detection starting from a single node
onestart_data <- onestart(adjmat=genmovnet_data, start.choice=2, stoch=FALSE)
```

## multistart function
Evaluate multiple nodes for invasion detection value 

```{r}
# Step 7: Evaluate multiple starts of invasion detection
multistart_data <- multistart(adjmat=genmovnet_data, stoch=FALSE)
```

## smartsurv function
Evaluates nodes for invaion detection value based on multiple realizations.

```{r}
# Step 8: Evaluate survival probabilities in the network
smartsurv_data <- smartsurv(adjmat=genmovnet_data, stoch=FALSE, nrealz=10)

print(smartsurv_data)
```

## smartsurv.weight function

includes probabilities for potential starting nodes for invasion other than that same as smartsurv

```{r}
# Step 9: Incorporate weights into the survival analysis
weights <- runif(nrow(genmovnet_data))  # Random weights for example
smartsurv_weight_data <- smartsurv.weight(ss.out=smartsurv_data, adjmat=genmovnet_data, wtvec=weights)

print(smartsurv_weight_data)
```

## spreadstep function
Generates a vector of status where each entry is the status of the node

```{r}
# Step 10: Simulate a spread step in the network
spreadstep_data <- spreadstep(amat=genmovnet_data, vect1=initvals_data)

print(spreadstep_data)
```

## makedec function
Generates plots based on coordinates of nodes that should adopt managment practices or new practices

```{r}
# Step 11: Make management decisions based on the communication vector
makedec_data <- makedec(comvec=sample(c(0, 1), size=100, replace=TRUE), 
                        geocoords4n= genlocs_data, 
                        readprobadoptvec4=FALSE, 
                        probadoptmean4=0.5, 
                        probadoptsd4=0.1, 
                        plotmp=TRUE)
```
## nsteps2

Determines bioentity establishment at each node

```{r}

# Step 12: Simulate multiple time steps in the network
nsteps_data <- ntsteps2(nsteps=3, 
                        infon=TRUE, 
                        geocoords3n=genlocs_data, 
                        vect1cn=rep(1, length(genlocs_data)/2), 
                        vect1dn=rep(0, length(genlocs_data)/2), 
                        readseam3=FALSE, 
                        seamdist3='powerlaw', 
                        seampla3=1, 
                        seamplb3=1, 
                        readbpam3=FALSE, 
                        bpamdist3='powerlaw', 
                        bpampla3=1, 
                        bpamplb3=1, 
                        probadoptmean3=0.5, 
                        probadoptsd3=0.1, 
                        probestabmean3=0.5, 
                        probestabsd3=0.1, 
                        maneffdir3='decrease_estab', 
                        maneffmean3n=0.5, 
                        maneffsd3n=0.1, 
                        readprobestabvec3=FALSE, 
                        readprobadoptvec3=FALSE, 
                        plotmpn=TRUE)
```
## reduce_nodes_to_focus

Reduces the number of nodes in the network to those which are linked to a set of key nodes

```{r}
# Step 13: Reduce the network to focus on key nodes
key_nodes <- c(1, 10)  
j <- genmovnet_data 

reduce_data <- reduce_nodes_to_focus(startmat=j, 
                                     maxdist=0, 
                                     keynodes=key_nodes, 
                                     tofrom='to')

# I HAVE NOT BEEN ABLE TO IMPLEMENT THE RUDUCED NETWORK I WOULD HAVE TO REDUCE MANY OTHER INFORMATIONS VECTORS USED IN INA AND MULTISAME
# FOR NOW REDUCED IS THE ENTIRE ADJACENY MATRIX

```

## INAscene function

Evaluates different scenarios in the network. For example the outputs are things like communication threshold and probability of managment adoptiona and bioentity establishment

```{r}
# Step 14: Perform an INA scenario
# For demonstration purposes; replace with your actual data currently random data from the matrix of genlocs
init_info_vector <- sample(c(0, 1), size=nrow(genlocs_data), replace=TRUE)
init_bio_vector <- sample(c(0, 1), size=nrow(genlocs_data), replace=TRUE)


INAscene_data <- INAscene(
  nreals=3, 
  ntimesteps=3, 
  doplot=TRUE, 
  readgeocoords=TRUE, 
  geocoords=genlocs_data,  
  numnodes=NA, 
  xrange=NA, 
  yrange=NA, 
  randgeo=FALSE, 
  readinitinfo=TRUE, 
  initinfo=init_info_vector,  
  readinitbio=TRUE, 
  initbio=init_bio_vector,  
  readseam=TRUE, 
  seam=j,     # For some reason you have to reaname the adjaceny matrix to i
  readbpam=FALSE, 
  bpamdist='random', 
  bpamrandp=0.1, 
  readprobadoptvec=FALSE, 
  probadoptmean=0.1, 
  probadoptsd=0.1, 
  readprobestabvec=FALSE, 
  probestabmean=0.1, 
  probestabsd=0.1, 
  maneffdir='decrease_estab', 
  maneffmean=0.5, 
  maneffsd=0.1, 
  usethreshman=FALSE
)
```

## multisame2 function

runs INA multiple times under different initial conditions

```{r}

# Step 15: Multi-scenario analysis using multsame2
multsame2_data <- multsame2(nreals2=10, 
                            ntimesteps2=3, 
                            usethreshman2=FALSE, 
                            readgeocoords2=TRUE, 
                            geocoords2=genlocs_data, 
                            maneffdir2='decrease_estab', 
                            maneffmean2=0.5, 
                            maneffsd2=0.1, 
                            maneffthresh2=0.5, 
                            sampeffort2=1, 
                            readinitinfo2=TRUE, 
                            initinfo2=init_info_vector, 
                            readinitbio2=TRUE, 
                            initbio2=init_bio_vector, 
                            readseam2=TRUE,                             
                            seam2=j, 
                            doplot2=TRUE,
                            readprobadoptvec2 = FALSE,
                            probadoptmean2 = 0.5,
                            probadoptsd2 = 0.1,
                            readbpam2 = FALSE,
                            bpamdist2 = 'random',
                            bpamrandp2 =0.5,
                            readprobestabvec2=FALSE,
                            probestabmean2 = 0.5,
                            probestabsd=0.5

                            )

# Better visualize the graphs also find a way to compare the diffrent visualizations


```
# layerplot and squish function
Plots the reduces bio and social networks with the added node information and revisualize the networks

```{r}

# Step 16: Visualization with layerplot and squish

#Need to Normalize data for layerplot
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}


j_normalized <- apply(j, 2, normalize)  
genlocs_data_normalized <- apply(genlocs_data, 2, normalize)  

upcoords <- genlocs_data_normalized
lowcoords <- genlocs_data_normalized


layerplot(upmat=j_normalized, lowmat=j_normalized, upcoords=upcoords, lowcoords=lowcoords, yflat=0.5, xover=0.5, htdif=1) # one color for bioentity and anotehr for social network

# Using squish for visualization
Z1 <- genlocs_data
plot(Z1, xlim=c(0,50), ylim=c(0,50), main="Original Coordinates")
Z2 <- squish(Z1)
plot(Z2, xlim=c(0,50), ylim=c(0,50), main="Squished Coordinates")




# TO-DO

# Improve visualizations and delivery of information from visualization

# Further investigate the use of these functions



```

